Elders PM Summary

### 2021-03-01 medvisit_data_cleaning.Rmd
data cleaning updates

---
title: "EldersAIR initial med visit data work"
author: "Ethan Walker"
date: "Started 20 Feb 2020, Updated 24 Feb 2020"
output: pdf_document
---

@@

# Initial load, format, save medical visit data
```{r}
## Box location: Update Box location once final data loaded
np_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NP/np_medicalvisit_20200220.xlsx"))

np_medvisit_new <- np_medvisit_initial %>% 
  mutate(home_winter_id = HomeWinterID,
         adult_id_num = AdultID,
         sp02 = as.numeric(Sp02),
         pr = as.numeric(PR),
         bp_sys = as.numeric(BPS),
         bp_dia = as.numeric(BPD),
         hr = as.numeric(HR),
         age = as.numeric(Age),
         gender = as.character(Gender),
         height_inches = as.numeric(Height),
         weight_lbs = as.numeric(Weight),
         fvc = as.numeric(PreFVC),
         fev1 = as.numeric(PreFEV),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fvc_pp = as.numeric(PredPreFVC),
         fev1_pp = as.numeric(PredPreFEV),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp),
         spiro_grade = as.character(PreGrade),
         spiro_maneuvers = as.numeric(PreManeuvers),
         spiro_status = as.character(DrStatus),
         spiro_result = as.character(DrResult),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  select(home_winter_id:area) %>% 
  arrange(home_winter_id)

# fix values after check spiro files
np_medvisit_new2 <- np_medvisit_new %>% 
  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
  mutate(sp02 = if_else(adult_id_num == 58 & sp02 == 80, -999, sp02)) %>% 
  mutate(pr = if_else(adult_id_num == 58 & pr == 80, -999, pr)) %>%
  mutate(bp_sys = if_else(adult_id_num == 58 & bp_sys == 80, -999, bp_sys)) %>%
  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
  mutate(fvc = if_else(adult_id_num == "47" & fvc == "1.01", 1.81, fvc),
         fev1 = if_else(adult_id_num == "55" & fev1 == "2.98", 1.98, fev1),
         fvc = if_else(adult_id_num == "104" & fvc == "1.2", 1.65, fvc),
         fev1 = if_else(adult_id_num == "104" & fev1 == "1.2", 1.6, fev1),
         fvc_pp = if_else(adult_id_num == "104" & fvc_pp == "42", 58, fvc_pp),
         fev1_pp = if_else(adult_id_num == "104" & fev1_pp == "56", 75, fev1_pp),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp))

## Box location: Update Box location once final data loaded
nn_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_medicalvisit_20200220.xlsx"))

@@

# fix values after check spiro files
nn_medvisit_new2 <- nn_medvisit_new %>% 
  mutate(bp_dia = if_else(bp_dia < 50, -999, bp_dia)) %>% 
  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
  mutate(hr = if_else(hr < 40, -999, hr)) %>% 
  mutate(fev1_fvc_ratio = if_else(fvc == -999, -999, fev1_fvc_ratio)) %>% 
  mutate(fev1_fvc_ratio_pp = if_else(fvc == -999, -999, fev1_fvc_ratio_pp)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
  replace_with_na(replace = list(bp_sys = c(9, 80))) %>% 
  mutate(age = if_else(adult_id_num == "47" & age == 17, 71, age)) %>% 
  mutate(fvc = if_else(adult_id_num == "87" & fvc == 2.65, 3.14, fvc),
         fev1 = if_else(adult_id_num == "87" & fev1 == 3.4, 2.65, fev1),
         fvc = if_else(adult_id_num == "77" & fvc == 2.19, 2.95, fvc),
         fev1 = if_else(adult_id_num == "77" & fev1 == 2.72, 2.22, fev1),
         fvc = if_else(adult_id_num == "77" & fvc == 0.81, 2.98, fvc),
         fev1 = if_else(adult_id_num == "77" & fev1 == 0.81, 2.38, fev1),
         fvc_pp = if_else(adult_id_num == "77" & fvc_pp == 57, 76, fvc_pp),
         fev1_pp = if_else(adult_id_num == "77" & fev1_pp == 49, 73, fev1_pp),
         fvc = if_else(adult_id_num == "55" & fvc == 1.13, 1.7, fvc),
         fev1 = if_else(adult_id_num == "55" & fev1 == 1.12, 1.18, fev1),
         fvc_pp = if_else(adult_id_num == "55" & fvc_pp == 0.35, 82, fvc_pp),
         fev1_pp = if_else(adult_id_num == "55" & fev1_pp == 2.31, 77, fev1_pp),
         fvc_pp = if_else(adult_id_num == "77" & fvc_pp == 21, 77, fvc_pp),
         fev1_pp = if_else(adult_id_num == "77" & fev1_pp == 27, 79, fev1_pp),
         fvc = if_else(adult_id_num == "79" & fvc == 1.59, 1.66, fvc),
         fev1 = if_else(adult_id_num == "79" & fev1 == 1.59, 1.19, fev1),
         fvc_pp = if_else(adult_id_num == "79" & fvc_pp == 89, 93, fvc_pp),
         fev1_pp = if_else(adult_id_num == "79" & fev1_pp == 123, 91, fev1_pp),
         fvc = if_else(adult_id_num == "61" & fvc == 3.07, 3.43, fvc),
         fvc_pp = if_else(fvc_pp < 1, fvc_pp*1000, fvc_pp),
         fev1_pp = if_else(fev1_pp < 1, fev1_pp*1000, fev1_pp),
         fvc_pp = if_else(fvc_pp < 10, (fvc/fvc_pp)*100, fvc_pp),
         fev1_pp = if_else(fev1_pp < 10, (fev1/fev1_pp)*100, fev1_pp),
         weight_lbs = if_else(adult_id_num == "61" & weight_lbs == 3.43, 127, weight_lbs),
         height_inches = if_else(height_inches == 5, 60, height_inches),
         bp_sys = if_else(adult_id_num == "9" & bp_sys == 94, 104, bp_sys),
         bp_dia = if_else(adult_id_num == "9" & bp_dia == 104, 94, bp_dia),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp))

elders_medvisit <- rbind(np_medvisit_new2, nn_medvisit_new2) %>% 
  select(-gender) %>% 
  mutate(bmi = (weight_lbs/(height_inches^2)*703),
         home_winter_id = as.factor(home_winter_id),
         adult_id_num = as.factor(adult_id_num)) %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id", "adult_id_num"))

#write_rds(elders_medvisit, "Output/elders_medvisit_clean.rds")

@@

write_csv(save_var_names, paste0(file_path, "Output/medvisit_var_names.csv"))
```

@@

>>>>

---
title: "EldersAIR initial med visit data work"
author: "Ethan Walker"
date: "Started 20 Feb 2020, Updated 1 March 2021"
output: pdf_document
---

@@

# Initial load, format, save medical visit data
```{r}
## Box location: Update Box location once final data loaded
npt_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NP/medicalvisit.xlsx"),
                                  na = "NULL")

npt_medvisit_new <- npt_medvisit_initial %>% 
  mutate(home_winter_id = HomeWinterID,
         adult_id_num = AdultID,
         sp02 = as.numeric(Sp02),
         pr = as.numeric(PR),
         bp_sys = as.numeric(BPS),
         bp_dia = as.numeric(BPD),
         hr = as.numeric(HR),
         age = as.numeric(Age),
         gender = as.character(Gender),
         height_inches = as.numeric(Height),
         weight_lbs = as.numeric(Weight),
         fvc = as.numeric(PreFVC),
         fev1 = as.numeric(PreFEV),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fvc_pp = as.numeric(PredPreFVC),
         fev1_pp = as.numeric(PredPreFEV),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp),
         spiro_grade = as.character(PreGrade),
         spiro_maneuvers = as.numeric(PreManeuvers),
         spiro_status = as.character(DrStatus),
         spiro_result = as.character(DrResult),
         sampling_visit = as.factor(SamplingVisit),
         area = "NPT") %>% 
  select(home_winter_id:area) %>% 
  arrange(home_winter_id)

# fix values after check spiro files
npt_medvisit_new2 <- npt_medvisit_new %>% 
  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
  mutate(sp02 = if_else(adult_id_num == 58 & sp02 == 80, -999, sp02)) %>% 
  mutate(pr = if_else(adult_id_num == 58 & pr == 80, -999, pr)) %>%
  mutate(bp_sys = if_else(adult_id_num == 58 & bp_sys == 80, -999, bp_sys)) %>%
  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
  mutate(fvc = if_else(adult_id_num == "47" & fvc == "1.01", 1.81, fvc),
         fev1 = if_else(adult_id_num == "55" & fev1 == "2.98", 1.98, fev1),
         fvc = if_else(adult_id_num == "104" & fvc == "1.2", 1.65, fvc),
         fev1 = if_else(adult_id_num == "104" & fev1 == "1.2", 1.6, fev1),
         fvc_pp = if_else(adult_id_num == "104" & fvc_pp == "42", 58, fvc_pp),
         fev1_pp = if_else(adult_id_num == "104" & fev1_pp == "56", 75, fev1_pp),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp))

## Box location: Update Box location once final data loaded
nn_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/medicalvisit.xlsx"),
                                  na = "NULL")

@@

# fix values after check spiro files
nn_medvisit_new2 <- nn_medvisit_new %>% 
  mutate(bp_dia = if_else(bp_dia < 50, -999, bp_dia)) %>% 
  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
  mutate(hr = if_else(hr < 40, -999, hr)) %>% 
  mutate(fev1_fvc_ratio = if_else(fvc == -999, -999, fev1_fvc_ratio)) %>% 
  mutate(fev1_fvc_ratio_pp = if_else(fvc == -999, -999, fev1_fvc_ratio_pp)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
  replace_with_na(replace = list(bp_sys = c(9, 80))) %>% 
  mutate(age = if_else(adult_id_num == "47" & age == 17, 71, age)) %>% 
  mutate(fvc = if_else(adult_id_num == "87" & fvc == 2.65, 3.14, fvc),
         fev1 = if_else(adult_id_num == "87" & fev1 == 3.4, 2.65, fev1),
         fvc = if_else(adult_id_num == "77" & fvc == 2.19, 2.95, fvc),
         fev1 = if_else(adult_id_num == "77" & fev1 == 2.72, 2.22, fev1),
         fvc = if_else(adult_id_num == "77" & fvc == 0.81, 2.98, fvc),
         fev1 = if_else(adult_id_num == "77" & fev1 == 0.81, 2.38, fev1),
         fvc_pp = if_else(adult_id_num == "77" & fvc_pp == 57, 76, fvc_pp),
         fev1_pp = if_else(adult_id_num == "77" & fev1_pp == 49, 73, fev1_pp),
         fvc = if_else(adult_id_num == "55" & fvc == 1.13, 1.7, fvc),
         fev1 = if_else(adult_id_num == "55" & fev1 == 1.12, 1.18, fev1),
         fvc_pp = if_else(adult_id_num == "55" & fvc_pp == 0.35, 82, fvc_pp),
         fev1_pp = if_else(adult_id_num == "55" & fev1_pp == 2.31, 77, fev1_pp),
         fvc_pp = if_else(adult_id_num == "77" & fvc_pp == 21, 77, fvc_pp),
         fev1_pp = if_else(adult_id_num == "77" & fev1_pp == 27, 79, fev1_pp),
         fvc = if_else(adult_id_num == "79" & fvc == 1.59, 1.66, fvc),
         fev1 = if_else(adult_id_num == "79" & fev1 == 1.59, 1.19, fev1),
         fvc_pp = if_else(adult_id_num == "79" & fvc_pp == 89, 93, fvc_pp),
         fev1_pp = if_else(adult_id_num == "79" & fev1_pp == 123, 91, fev1_pp),
         fvc = if_else(adult_id_num == "61" & fvc == 3.07, 3.43, fvc),
         fvc_pp = if_else(fvc_pp < 1, fvc_pp*1000, fvc_pp),
         fev1_pp = if_else(fev1_pp < 1, fev1_pp*1000, fev1_pp),
         fvc_pp = if_else(fvc_pp < 10, (fvc/fvc_pp)*100, fvc_pp),
         fev1_pp = if_else(fev1_pp < 10, (fev1/fev1_pp)*100, fev1_pp),
         weight_lbs = if_else(adult_id_num == "61" & weight_lbs == 3.43, 127, weight_lbs),
         height_inches = if_else(height_inches == 5, 60, height_inches),
         height_inches = if_else(height_inches == 79.8, 60, height_inches),
         bp_sys = if_else(adult_id_num == "9" & bp_sys == 94, 104, bp_sys),
         bp_dia = if_else(adult_id_num == "9" & bp_dia == 104, 94, bp_dia),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp))

elders_medvisit <- rbind(npt_medvisit_new2, nn_medvisit_new2) %>% 
  select(-gender) %>% 
  mutate(bmi = (weight_lbs/(height_inches^2)*703),
         home_winter_id = as.factor(home_winter_id),
         adult_id_num = as.factor(adult_id_num)) %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id", "adult_id_num")) %>% 
  arrange(area, adult_id_num, winter_id, sampling_visit) %>% 
  select(area, home, home_id_num, home_winter_id, winter_id, treatment, 
         adult_id_char, adult_id_num, sampling_visit,
         sp02:weight_lbs, bmi, fvc:spiro_result)

write_rds(elders_medvisit, paste0(file_path, "Output/elders_medvisit_clean.rds"))

@@

#write_csv(save_var_names, paste0(file_path, "Output/medvisit_var_names.csv"))
```

@@

# Further data checking/cleaning
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds"))

medvisit_data_checks <- elders_medvisit %>% 
  mutate(bp_check = if_else(bp_dia >= bp_sys, 1, 0), # no instances of dbp >= sbp
         hr_check = if_else(pr-hr >= abs(10), 1, 0), # 53 instances of pr/hr diff >= 10
         fvc_check = if_else(fev1 >= fvc, 1, 0)) %>% # 2 instances of fvc = fev1; none where fev1 > fvc
  filter(fvc_check == 1)

table(elders_medvisit$spiro_grade)
table(elders_medvisit$spiro_status)
table(elders_medvisit$spiro_result)
```

### 2021-03-01 pm_summary_dustrak.Rmd
data cleaning updates

winter2_summary <- elders_pm_winter %>% 
  left_join(elders_kw_report_clean, by = c("area", "home",
                                           "treatment")) %>% 
  mutate(treatment = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type = if_else(treatment == "Filter", filter_type, treatment),
         treatment_filter_type = factor(treatment_filter_type,
                                        levels = c("Placebo", "Education",
                                                   "Large Filtrete", "Winix"))) %>% 
  #filter(treatment_filter_type != "Filter" & treatment_filter_type != "Placebo") %>% 
  filter(!is.na(treatment_filter_type)) %>% 
  #filter(winter_id.x == 2) %>% 
  group_by(winter_id.x, treatment_filter_type) %>% 
  summarize("Mean PM" = mean(pm_mean_winter, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(pm_mean_winter, na.rm = TRUE),
            "Min PM" = min(pm_mean_winter, na.rm = TRUE),
            "Median PM" = median(pm_mean_winter, na.rm = TRUE),
            "Max PM" = max(pm_mean_winter, na.rm = TRUE))
winter2_summary

>>>>

winter2_summary <- elders_pm_winter %>% 
  left_join(elders_kw_report_clean, by = c("area", "home",
                                           "treatment")) %>% 
  mutate(treatment = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type = if_else(treatment == "Filter", filter_type, treatment),
         treatment_filter_type = factor(treatment_filter_type,
                                        levels = c("Placebo", "Education",
                                                   "Large Filtrete", "Winix"))) %>% 
  #filter(treatment_filter_type != "Filter" & treatment_filter_type != "Placebo") %>% 
  filter(!is.na(treatment_filter_type)) %>% 
  #filter(winter_id.x == 2) %>% 
  group_by(winter_id.x, treatment) %>% 
  summarize("Mean PM" = mean(pm_mean_winter, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(pm_mean_winter, na.rm = TRUE),
            "Min PM" = min(pm_mean_winter, na.rm = TRUE),
            "Median PM" = median(pm_mean_winter, na.rm = TRUE),
            "Max PM" = max(pm_mean_winter, na.rm = TRUE))
winter2_summary
